|

### 2023-04-10 paper/01-introduction.Rmd
fix knitting issues

---
output:
  html_document: default
  pdf_document: default
---
# Introduction {#sec:introduction}

@@

In recent years, tracking data have replaced traditional box-score statistics and play-by-play data as the state-of-the-art tool in sports analytics.Numerous sports have been collecting and releasing data on player and ball locations on the playing surface over the course of a game.

This spatio-temporal multi-resolution source of data have provided exceptional opportunities for researchers to perform advanced studies at a more complex level to deepen the understanding of different sports.

For complete surveys on how tracking data have transformed sports analytics, see @Macdonald2020Recreating, @Baumer2023Big, and @Kovalchik2023Player.

@@

## Previous Pass Rush Metrics
Officially tracked by the NFL since 1982, the most well known statistic for evaluating defensive linemen on pass rushing plays is the sack, which occurs when a defensive player tackles the quarterback behind the line of scrimmage before the quarterback is able to throw a pass.  Sacks represent a simple binary measure of success for a rusher on a pass play.  However, there are many intermediate outcomes for a rusher that can also be considered successes that don't result in a sack.  Measures such a hits and hurries are collected by various outlets where hits counts the number of times a QB is hit while passing and hurries counts how often a QB has to throw the ball earlier than they would have liked to.

@@

\begin{table}
\caption{A summary of previously-existing pass rush metrics. \label{tab:metrics}}
\begin{center}
\begin{tabular}{ll}
\textbf{Metric} & \textbf{Description}\\\hline
Sacks & A defensive player tackles the quarterback behind the line of scrimmage before the quarterback is able to throw a pass\\
Hits & A defensive player tackles the quarterback behind the line of scrimmage after the quarterback is able to throw a pass.   \\
Hurries & A defensive player pressures the quarterback behind the line of scrimmage such that the quarterback must release the ball prior sooner than they wanted to. \\
QB Pressures & A sack, hit, or hurry.  \\
Pass-Rush Productivity (PRP) & $\frac{\frac{(\# Hurries + \# Hits)}{2} + \# Sacks}{\#PassRush Snaps}$\\
Time In Pocket (TIP) & The time from snap to throw or pocket collapse for a quarterback (https://www.pff.com/news/pro-pff-signature-statistics-a-glossary#:~:text=Time%20in%20Pocket,time%20until%20a%20quarterback%20scrambles)\\
Pass Rush Win Rate (PRWR) & Rate at which pass rusher beats pass block within 2.5 seconds after snap (https://www.espn.com/nfl/story/_/id/34536376/2022-nfl-pass-rushing-run-stopping-blocking-leaderboard-win-rate-rankings-top-players-teams)\\\hline
\end{tabular}
\end{center}
\end{table}

## Previous Pass-Rush Measure in academic literaure
The academic literature on the performance of linemen, either offensive or defensive in football is scant. @AlamarGould2008 studied the the relationship between pass completion rate and performance of offensive linemen in terms of how successful they were at blocking.  The data for this study was collecting essentially by hand by watching and charting the performance of each lineman on a play.  This was done using games from the beginning 2007 season for seven teams and their first three games each.  The authors note that there is not much research on linemen saying "One of the principal reasons for this lack of study is the absence of publicly available data that is needed in order to track the performance of a specific lineman."

@@

## Contributions of STRAIN
All the aforementioned methods suffered from not having the data that was actually necessary to develop a meaningful metric for linemen in football, either offensive or defensive.  Now though, thanks to the 2023 Big Data Bowl, we have a publicly available tracking data not only for the linemen, but for every player on the field!  With this data performance of offensive and defensive linemen can be meaningfully assessed, though we focus solely on the performance of defensive linemen here.  This manuscript presents a metric we refer to as STRAIN based on a concept of the same name from materials science.  STRAIN gives a continuous measure of pressure for each of the defensive linemen over the course of an entire play allowing for the measurement of the level of success of the pass rush even on plays that don't result in an easily observed outcome like a sack, hit, or hurry.  We view this as a major step forward in the accurate evaluation of defensive linemen performance.  Further, we note than while we focus exclusively on defensive linemen here, the proposed metric STRAIN is relatively easily extended to the evaluation of offensive linemen performance.

>>>>

---
output:
  html_document: default
  pdf_document: default
---

# Introduction {#sec:introduction}

@@

In recent years, tracking data have replaced traditional box-score statistics and play-by-play data as the state-of-the-art tool in sports analytics.
Numerous sports have been collecting and releasing data on player and ball locations on the playing surface over the course of a game.
This spatio-temporal multi-resolution source of data have provided exceptional opportunities for researchers to perform advanced studies at a more complex level to deepen the understanding of different sports.
For complete surveys on how tracking data have transformed sports analytics, see @Macdonald2020Recreating, @Baumer2023Big, and @Kovalchik2023Player.

@@

## Previous Pass Rush Metrics

Officially tracked by the NFL since 1982, the most well known statistic for evaluating defensive linemen on pass rushing plays is the sack, which occurs when a defensive player tackles the quarterback behind the line of scrimmage before the quarterback is able to throw a pass.  Sacks represent a simple binary measure of success for a rusher on a pass play.  However, there are many intermediate outcomes for a rusher that can also be considered successes that don't result in a sack.  Measures such a hits and hurries are collected by various outlets where hits counts the number of times a QB is hit while passing and hurries counts how often a QB has to throw the ball earlier than they would have liked to.

@@

\begin{table}
\caption{A summary of previously-existing pass rush metrics. \label{tab:metrics}}
\begin{center}
\begin{tabular}{ll}
\textbf{Metric} & \textbf{Description}\\\hline
Sacks & A defensive player tackles the quarterback behind the line of scrimmage before the quarterback is able to throw a pass \\
Hits & A defensive player tackles the quarterback behind the line of scrimmage after the quarterback is able to throw a pass. \\
Hurries & A defensive player pressures the quarterback behind the line of scrimmage such that the quarterback must release the ball prior sooner than they wanted to. \\
QB Pressures & A sack, hit, or hurry. \\
Pass-Rush Productivity (PRP) & $\frac{\frac{(\# Hurries + \# Hits)}{2} + \# Sacks}{\#PassRush Snaps}$\\
Time In Pocket (TIP) & The time from snap to throw or pocket collapse for a quarterback \\
Pass Rush Win Rate (PRWR) & Rate at which pass rusher beats pass block within 2.5 seconds after snap\\\hline
\end{tabular}
\end{center}
\end{table}

(https://www.pff.com/news/pro-pff-signature-statistics-a-glossary#:~:text=Time%20in%20Pocket,time%20until%20a%20quarterback%20scrambles)

## Previous Pass-Rush Measure in academic literaure

The academic literature on the performance of linemen, either offensive or defensive in football is scant. @AlamarGould2008 studied the the relationship between pass completion rate and performance of offensive linemen in terms of how successful they were at blocking.  The data for this study was collecting essentially by hand by watching and charting the performance of each lineman on a play.  This was done using games from the beginning 2007 season for seven teams and their first three games each.  The authors note that there is not much research on linemen saying "One of the principal reasons for this lack of study is the absence of publicly available data that is needed in order to track the performance of a specific lineman."

@@

## Contributions of STRAIN

All the aforementioned methods suffered from not having the data that was actually necessary to develop a meaningful metric for linemen in football, either offensive or defensive.  Now though, thanks to the 2023 Big Data Bowl, we have a publicly available tracking data not only for the linemen, but for every player on the field!  With this data performance of offensive and defensive linemen can be meaningfully assessed, though we focus solely on the performance of defensive linemen here.  This manuscript presents a metric we refer to as STRAIN based on a concept of the same name from materials science.  STRAIN gives a continuous measure of pressure for each of the defensive linemen over the course of an entire play allowing for the measurement of the level of success of the pass rush even on plays that don't result in an easily observed outcome like a sack, hit, or hurry.  We view this as a major step forward in the accurate evaluation of defensive linemen performance.  Further, we note than while we focus exclusively on defensive linemen here, the proposed metric STRAIN is relatively easily extended to the evaluation of offensive linemen performance.
